The forestecology R package for fitting and assessing neighborhood models of the effect of interspecific competition on the growth of trees

### 2021-08-06 paper/paper.Rmd
results for appedices 3+4 need to edit text

# Set output width
options(
  width = 72, 
  # Decimal precision
  digits = 3, 
  # Number of rows for tibble printing
  tibble.print_max = 5, 
  tibble.print_min = 5
)

# TODO:
setwd("~/Documents/forestecology/forestecology/paper/")

# TODO: set eval=TRUE for Appendix code blocks
```

@@

>>>>

# Set output width
options(
  width = 72, 
  # Decimal precision
  digits = 3, 
  # Number of rows for tibble printing
  tibble.print_max = 5, 
  tibble.print_min = 5
)
```

@@

# Appendix 3: Compare predictor variables {#compare-pred-var}

```{r compar-pred-var}
census_2013_scbi <- read_csv("scbi.stem2.csv") %>%
  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
  mutate(
    date = mdy(date),
    dbh = as.numeric(dbh)/10
  ) %>%
  filter(gx < 300, between(gy, 300, 600))

census_2018_scbi <- read_csv("scbi.stem3.csv") %>%
  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
  mutate(
    date = mdy(date),
    dbh = as.numeric(dbh)/10
  ) %>%
  filter(gx < 300, between(gy, 300, 600))

growth_scbi <-
  compute_growth(
    census_1 = census_2013_scbi,
    census_2 = census_2018_scbi %>% filter(!str_detect(codes, "R")),
    id = "stemID"
  ) %>%
  left_join(sp_info, by = "sp") %>% 
  mutate(sp = as.factor(sp),
        basal_area = 0.0001 * pi * (dbh1 / 2)^2,
         agb = get_biomass(
                  dbh = dbh1,
                  genus = genus,
                  species = species,
                  coords = c(-78.2, 38.9)
    ))

n_fold <- 4
fold1 <- cbind(c(0, 150, 150, 0), c(300, 300, 450, 450))
fold2 <- cbind(c(150, 300, 300, 150), c(300, 300, 450, 450))
fold3 <- cbind(c(0, 150, 150, 0), c(450, 450, 600, 600))
fold4 <- cbind(c(150, 300, 300, 150), c(450, 450, 600, 600))

comp_dist <- 7.5

focal_vs_comp_ba <- growth_scbi %>%
    create_focal_vs_comp(comp_dist = comp_dist, 
                         blocks = blocks_scbi, 
                         id = "stemID", 
                         comp_x_var = "basal_area") %>%
    run_cv(comp_dist = comp_dist, blocks = blocks_scbi)

focal_vs_comp_agb <- growth_scbi %>%
    create_focal_vs_comp(comp_dist = comp_dist, 
                         blocks = blocks_scbi, 
                         id = "stemID", 
                         comp_x_var = "agb") %>%
    run_cv(comp_dist = comp_dist, blocks = blocks_scbi)

  focal_vs_comp_ba %>%
    rmse(truth = growth, estimate = growth_hat) %>%
    pull(.estimate)

  focal_vs_comp_agb %>%
    rmse(truth = growth, estimate = growth_hat) %>%
    pull(.estimate)

```

# Appendix 4: Compare grouping variables {#compare-group-var}

Creative use of the grouping variable allows for tests of what best explains competitive interactions. Here we compare whether species idenity or canopy position of the species best explains.

```{r compare-group-var}

census_2013_scbi <- read_csv("scbi.stem2.csv") %>%
  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
  mutate(
    date = mdy(date),
    dbh = as.numeric(dbh)/10
  ) %>%
  filter(gx < 300, between(gy, 300, 600))

census_2018_scbi <- read_csv("scbi.stem3.csv") %>%
  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
  mutate(
    date = mdy(date),
    dbh = as.numeric(dbh)/10
  ) %>%
  filter(gx < 300, between(gy, 300, 600))

growth_scbi_sp <-
  compute_growth(
    census_1 = census_2013_scbi,
    census_2 = census_2018_scbi %>% filter(!str_detect(codes, "R")),
    id = "stemID"
  ) %>%
  mutate(sp = as.factor(sp),
        basal_area = 0.0001 * pi * (dbh1 / 2)^2 )

growth_scbi_can_pos <-
  compute_growth(
    census_1 = census_2013_scbi,
    census_2 = census_2018_scbi %>% filter(!str_detect(codes, "R")),
    id = "stemID"
  ) %>%
  left_join(sp_info, by = "sp") %>% 
  mutate(
    canopy_position = str_replace(canopy_position, ' ', '_'),
    canopy_position = str_replace(canopy_position, ',', ''),
    canopy_position = ifelse(is.na(canopy_position), 'shrub_layer', canopy_position),
    sp = as.factor(canopy_position),
    basal_area = 0.0001 * pi * (dbh1 / 2)^2 )

n_fold <- 4
fold1 <- cbind(c(0, 150, 150, 0), c(300, 300, 450, 450))
fold2 <- cbind(c(150, 300, 300, 150), c(300, 300, 450, 450))
fold3 <- cbind(c(0, 150, 150, 0), c(450, 450, 600, 600))
fold4 <- cbind(c(150, 300, 300, 150), c(450, 450, 600, 600))

growth_scbi_sp <- growth_scbi_sp %>%
  mutate(foldID = spatial_block_scbi$foldID %>% factor())
growth_scbi_can_pos <- growth_scbi_can_pos %>%
  mutate(foldID = spatial_block_scbi$foldID %>% factor())

comp_dist <- 7.5

growth_scbi_sp <- growth_scbi_sp %>%
    add_buffer_variable(size = comp_dist, region = study_region_scbi)
growth_scbi_can_pos <- growth_scbi_can_pos %>%
    add_buffer_variable(size = comp_dist, region = study_region_scbi)

focal_vs_comp_sp <- growth_scbi_sp %>%
    create_focal_vs_comp(comp_dist = comp_dist, 
                         blocks = blocks_scbi, 
                         id = "stemID", 
                         comp_x_var = "basal_area") %>%
    run_cv(comp_dist = comp_dist, blocks = blocks_scbi)

focal_vs_comp_can_pos <- growth_scbi_can_pos %>%
    create_focal_vs_comp(comp_dist = comp_dist, 
                         blocks = blocks_scbi, 
                         id = "stemID", 
                         comp_x_var = "basal_area") %>%
    run_cv(comp_dist = comp_dist, blocks = blocks_scbi)

  focal_vs_comp_sp %>%
    rmse(truth = growth, estimate = growth_hat) %>%
    pull(.estimate)

  focal_vs_comp_can_pos %>%
    rmse(truth = growth, estimate = growth_hat) %>%
    pull(.estimate)
```

We find that species identity has a lower RMSE, so does a better job. But we can still plot the competition posteriors for the canopoy position groupings.

```{r can-pos-comp}

  fit_mod_can_pos <- growth_scbi_can_pos %>%
    create_focal_vs_comp(comp_dist = comp_dist, 
                         blocks = blocks_scbi, 
                         id = "stemID", 
                         comp_x_var = "basal_area")  %>%
    comp_bayes_lm(prior_param = NULL)

  autoplot(fit_mod_can_pos, type = "competition")

```
